MG7: Configurable and scalable 16S metagenomics data analysis -- new methods optimized for massive cloud computing

### 2015-09-13 docs/paper/out/paper.tex
improve and expand the idea of data analysis as software -- paper

General approach. An analysis is defined as a software project. It can
evolve in the same way. We can run the analysis in a test phase, review
configuration and changes, etc. Key advantages of this approach are

\begin{itemize}
\tightlist
\item
  \textbf{Reproducibility} the same analysis can be run again with
  exactly the same configuration in a trivial way.
\item
  \textbf{Versioning} The analysis is a software project so it goes
  through the same stages, there can be different versions, stable
  releases, etc.
\item
  \textbf{Reuse} we can build standard configurations on top of this and
  reuse them for subsequent data analysis.
\item
  \textbf{Decoupling} We can start working on the analysis
  specification, without any need for data in a much easier way.
\item
  \textbf{Expresiveness and safety} choose only from valid Illumina read
  types, build default FLASH command based on that, \ldots{}
\end{itemize}

>>>>

MG7 proposes to define and work with a particular data analysis task as
a software project, using Scala. The idea is that \emph{everything}:
data description, their location, configuration parameters, the
infrastructure used, \ldots{} should be expressed as Scala code, and
treated in the same way as any (well-managed) software project. This
includes, among other things, using version control systems
(\texttt{git} in our case), writing tests, making stable releases
following \href{http://semver.org/}{semantic versioning} or publishing
artifacts to a repository.

What we see as key advantages of this approach (when coupled with
compile-time specification and checking), are

\begin{itemize}
\tightlist
\item
  \textbf{Reproducibility} the same analysis can be run again with
  exactly the same configuration in a trivial way.
\item
  \textbf{Versioning} as in any software project, there can be different
  versions, stable releases, etc.
\item
  \textbf{Reuse} we can build standard configurations on top of this and
  reuse them for subsequent data analysis. A particular data analysis
  \emph{task} can be used as a \emph{library} in further analysis.
\item
  \textbf{Decoupling} We can start working on the analysis
  specification, without any need for available data in a much easier
  way.
\item
  \textbf{Documentation} We can take advantage of all the effort put
  into software documentation tools and practices, such as in our case
  Scaladoc or literate programming. As documentation, analysis processes
  and data specification live together in the files, it is much easier
  to keep coherence between them.
\item
  \textbf{Expresiveness and safety} For example in our case we can
  choose only from valid Illumina read types, and then build a default
  FLASH command based on that. The output locations, being declared
  statically, are also available for use in further analysis.
\end{itemize}
